### 2010-02-25 trunk/deliverables/D2_2_Ontology/properties.tex

\subsubsection*{Property: \texttt{fgo:contains}}
\label{subs:contains}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & contains \\ \hline
\textcolor{white}{\textbf{description}} & Many kinds of components in FAST can contain other components: 
screenflows contain screens, screens contain forms or form elements, etc.
It points to a resource reference to give a unique `id' to each component
within the resource. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:from}}
\label{subs:from}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & from \\ \hline
\textcolor{white}{\textbf{description}} & This property defines the starting point of a pipe within a screen. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasAction}}
\label{subs:hasAction}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has action \\ \hline
\textcolor{white}{\textbf{description}} & This property indicates which actions are associated and can be perfomed within 
a screen component. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Action}}{subs:Action} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasCode}}
\label{subs:hasCode}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has code \\ \hline
\textcolor{white}{\textbf{description}} & URL of the executable code of a particular building block (currently screens or screen components). \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:WithCode}}{subs:WithCode} \\ \hline
\textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasCopy}}
\label{subs:hasCopy}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has copy \\ \hline
\textcolor{white}{\textbf{description}} & Links the template of a resource as used in the palette or catalogue of available resources to its copy, as used in a particular screen flow. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasForm}}
\label{subs:hasForm}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has form \\ \hline
\textcolor{white}{\textbf{description}} & If a screen is defined declaratively, this property links it to its form (i.e., its visual user interface). \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Screen}}{subs:Screen} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Form}}{subs:Form} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasFormElement}}
\label{subs:hasFormElement}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has form element \\ \hline
\textcolor{white}{\textbf{description}} & If a form is defined declaratively, its elements are linked to it with this property. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Form}}{subs:Form} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:FormElement}}{subs:FormElement} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasIcon}}
\label{subs:hasIcon}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has icon \\ \hline
\textcolor{white}{\textbf{description}} & A small graphical representation of any FAST component or sub-component. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
\textcolor{white}{\textbf{range}} & \texttt{foaf:Image} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasLanguage}}
\label{subs:hasLanguage}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has language string \\ \hline
\textcolor{white}{\textbf{description}} & This property is the language the library is written in. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasLibrary}}
\label{subs:hasLibrary}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has library \\ \hline
\textcolor{white}{\textbf{description}} & This property indicates which libraries are used by a screen component. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:WithCode}}{subs:WithCode} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasOperator}}
\label{subs:hasOperator}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has operator \\ \hline
\textcolor{white}{\textbf{description}} & If a screen is defined declaratively, this property links it to an operator it might contain. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Screen}}{subs:Screen} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Operator}}{subs:Operator} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasPattern}}
\label{subs:hasPattern}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has pattern \\ \hline
\textcolor{white}{\textbf{description}} & This property links the abstract representation of a fact (i.e., an RDF triple) in FAST to a graph resource containing the actual triple. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
\textcolor{white}{\textbf{range}} & \texttt{rdfs:Resource} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasPatternString}}
\label{subs:hasPatternString}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has pattern string \\ \hline
\textcolor{white}{\textbf{description}} & This property links to the textual representation of a fact (i.e., an RDF triple), e.g., in N3. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Fact}}{subs:Fact} \\ \hline
\textcolor{white}{\textbf{range}} & \texttt{xsd:Condition} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasPostCondition}}
\label{subs:hasPostCondition}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has post-condition \\ \hline
\textcolor{white}{\textbf{description}} & This property links certain type of resources to its post-condition, 
i.e., the facts that are produced once the screen or screenflow has been 
executed. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:WithPostConditions}}{subs:WithPostConditions} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasPreCondition}}
\label{subs:hasPreCondition}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has pre-condition \\ \hline
\textcolor{white}{\textbf{description}} & This property links certain type of resources to its pre-condition, i.e., 
the facts that need to be fulfilled in order for this screen or screenflow to be 
reachable. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:WithPreConditions}}{subs:WithPreConditions} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasResource}}
\label{subs:hasResource}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has resource \\ \hline
\textcolor{white}{\textbf{description}} & If a screen is defined declaratively, this property links it to its service resource (i.e., its backend). \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Screen}}{subs:Screen} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasScreenshot}}
\label{subs:hasScreenshot}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has screenshot \\ \hline
\textcolor{white}{\textbf{description}} & An image which shows a particular screen or screenflow in action, to 
aid users in deciding which screen or screenflow to choose out of many. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
\textcolor{white}{\textbf{range}} & \texttt{foaf:Image} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasTemplate}}
\label{subs:hasTemplate}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has template \\ \hline
\textcolor{white}{\textbf{description}} & Links the copy of a resource as used in a particular screen flow to its template, as used in the palette or catalogue of available resources. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:hasTrigger}}
\label{subs:hasTrigger}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has trigger string \\ \hline
\textcolor{white}{\textbf{description}} & This property links a buildiong block to a Trigger. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:isPositive}}
\label{subs:isPositive}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & is positive \\ \hline
\textcolor{white}{\textbf{description}} & Facts can be set to a specific scope: design time, execution time,
or both of them. This property will define when they have to be taken into account by the
inference engine or reasoner. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
\textcolor{white}{\textbf{range}} & \texttt{xsd:boolean} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:to}}
\label{subs:to}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & to \\ \hline
\textcolor{white}{\textbf{description}} & This property defines the end point of a pipe within a screen. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
\end{tabular}
\subsubsection*{Property: \texttt{fgo:uses}}
\label{subs:uses}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & uses \\ \hline
\textcolor{white}{\textbf{description}} & This property indicates concepts used within a building block, without being a 
pre- or post-condition. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
\textcolor{white}{\textbf{range}} & \texttt{rdfs:Resource} \\ \hline
\end{tabular}