### 2010-06-14 trunk/papers/icsoc2010/paper.tex
split it into different files
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@844 d8b0c59b-6544-0410-94d9-83a5658a058e

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{url}
\urldef{\mailsa}\path|{ismael.rivera, knud.moeller}@deri.org| 
\urldef{\mailsb}\path|zuendorf@cs.uni-kassel.de|
%\urldef{\mailsc}\path|erika.siebert-cole, peter.strasser, lncs}@springer.com|    
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\usepackage{amssymb}

\newenvironment{listing}
{\begin{list}{}{\setlength{\leftmargin}{1em}}\item\scriptsize\bfseries}
{\end{list}}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{An improved platform for Web service wrapping, discovery and consumption}

% a short form should be given in case it is too long for the running head
\titlerunning{Enhancing the discovery and consumption of Web services}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Ismael Rivera\and Knud M\"oller\and Albert Z\"undorf}
%
\authorrunning{Ismael Rivera\and Knud M\"oller\and Albert Z\"undorf}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{DERI, National University of Ireland, Galway\\
University of Kassel, Germany\\
\mailsa\\
\mailsb\\
%\mailsc\\
%\url{http://www.springer.com/lncs}
}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

\toctitle{Lecture Notes in Computer Science}
\tocauthor{Authors' Instructions}
\maketitle

\input{paper_abstract}

\input{paper_sec_introduction}

\input{paper_sec_related_work}

\input{paper_sec_wrapping}

\input{paper_sec_catalogue}

\input{paper_sec_use_case}

\input{paper_sec_conclusions}

\subsubsection*{Acknowledgments.} This work is supported in part by the European Commission under the first call of its Seventh Framework Program (FAST STREP Project, grant INFSO-ICT-216048) and in part by Science Foundation Ireland under Grant No. SFI/08/CE/I1380 (L\'ion-2).

\bibliographystyle{apalike}
\bibliography{paper}

\section*{Checklist of Items to be Sent to Volume Editors}
Here is a checklist of everything the volume editor requires from you:

\begin{itemize}
\settowidth{\leftmargin}{{\Large$\square$}}\advance\leftmargin\labelsep
\itemsep8pt\relax
\renewcommand\labelitemi{{\lower1.5pt\hbox{\Large$\square$}}}

\item The final \LaTeX{} source files
\item A final PDF file
\item A copyright form, signed by one author on behalf of all of the
authors of the paper.
\item A readme giving the name and email address of the
corresponding author.
\end{itemize}
\end{document}