### 2010-06-14 trunk/papers/icsoc2010/paper_sec_conclusions.tex

\section{Conclusions and future work}

Future work:

Talk about wrapping WSDL-based services

Talk about WSMO-lite services

Use JSONRequest to replace the AJAX XMLHttpRequest and its problems (cross-domain calls), so there's no need for a Proxy.