### 2010-06-15 trunk/papers/icsoc2010/paper_abstract.tex
last changes everywhere
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@876 d8b0c59b-6544-0410-94d9-83a5658a058e

%!TEX root = paper.tex

\begin{abstract}
B2B systems integration was revolutionised by the introduction of web services,
transforming the way enterprise systems communicate and strengthening the  relationships with providers and customers. The main advantage seen by companies was an increase of operational efficiencies, and reduction of costs. 
In this scenario, highly qualified software developers are responsible for the integration of services with other systems, through the analysis of formal service description or human-readable specifications. 
However, this model fails when targeting the long tail of enterprise software demand, the end-users. Discovery and consumption of web services is far from a straight-forward task for an end-user. This means that potential users trying to create task-specific applications are left out. 
This paper presents an approach to facilitate the discovery and consumption of business web services by end-users, closing the gap between the two. The approach includes: (a) a method to generate ready-to-use web service wrappers, and (b) a catalogue which users can browse to search for web services fitting their needs. 
% These web services wrappers are described by a model created, although highly influenced by well-known web services modelling standards. 
\keywords{web services, end-users, discovery, consumption, linked data}
\end{abstract}