### 2010-09-28 trunk/papers/webist2011/paper_ssec_discovery.tex
added discovery section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@983 d8b0c59b-6544-0410-94d9-83a5658a058e

%!TEX root = paper.tex

\subsection{Discovery mechanisms} % (fold)
\label{ssec:discovery}

The platform presented in this paper is intended to support web service discovery and recommendation capabilities.
As explained in detail in Section~\ref{ssec:ontology}, there are different types of building block which can be modelled
using the platform, leading to a necessity for different kind of discovery or recommendation mechanisms, depending
on the level of composition to deal with: screen-flow or screen. In the screen-flow composition, the smallest
functional unit is the screen, hance a screen-flow is composed by interconnecting a number of screens between 
each other. In fact, this composition can be applied to any sort of building block since they all share the
same structure: a set of pre- and post-conditions which make possible the communication with other building blocks.
That said, technically the algorithms presented in this section could be used for the discovery and composition of
any type of building block stored in the platform.

At the screen-flow design level, the composition is made using a set of screens, and the pre- and post-conditions 
of the screen-flow itself, used as entry and exit points. A screen is only reachable when its pre-conditions are satisfied
as well, meaning that the post-conditions of another screen, or the pre-conditions of the screen-flow are compatible 
with them.

The main goal of the discovery process is to help the user to make the screen-flow executable, in other words, to make
all the screens within a screen-flow reachable. To do that, the platform offers two mechanisms to find and recommend
screens, previously stored in the building block base (the catalogue), which will make the screen-flow executable: a
simple discovery based on pre- and postconditions, and a multi-step discovery or planning mechanism. Like any other
search engine or recommender system, the results obtained in the discovery process need to rank before being presented
to the user. The ranking algorithm is covered in Section~\ref{sssec:ranking}.

\subsubsection{Simple discovery based on pre- and post-conditions}
\label{sssec:simple_discovery}

At a particular state of a composition, ie. at screen-flow design, several of the building blocks used to compose it, 
screens in the case of a screen-flow, might be unreachable, therefore there must be pre-conditions not satisfied. The
platform will assist the process by recommending building blocks which will satisfy these pre-conditions. In a nutshell,
it collects the pre-conditions of all the unreachable building blocks, used as a graph pattern (see Section~\ref{ssec:ontology})
in order to be matched against the post-conditions of any building block. In the following scenario, there are two 
screens: \emph{s1} and \emph{s2}. \emph{s1} has as a pre-condition: ``there exists a search criteria'', and as a 
post-condition: ``there exists a item''. \emph{s2} has just a pre-condition stating: ``there exists a search criteria''.
The same example formally defined:

\begin{listing}
\begin{verbatim}
:G1 { :s1 a fgo:Screen .      
      :s1 fgo:hasPrecondition c1 .
      :s1 fgo:hasPostcondition c2 .
      :c1 fgo:hasPattern GC1 .
      :c2 fgo:hasPattern GC2 .
      :s2 a fgo:Screen .      
      :s2 fgo:hasPrecondition c3 .
      :c3 fgo:hasPattern GC3 }
:GC1 { _:x a amazon:SearchCriteria }
:GC2 { _:x a amazon:Item }
:GC3 { _:x a amazon:SearchCriteria }
\end{verbatim}
\label{lis:discovery_example}
\end{listing}

The algorithm will construct a SPARQL query to retrieve building blocks satisfying the pre-conditions \emph{c1} and \emph{c3}. 
It would look something like:

\begin{listing}
\begin{verbatim}
PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
PREFIX fgo: <http://purl.oclc.org/fast/ontology/gadget#>
PREFIX amazon:<http://aws.amazon.com/AWSECommerceService#>

SELECT DISTINCT ?bb 
WHERE { 
  ?bb rdf:type fgo:Screen . 
  {
    {
      ?bb fgo:hasPostCondition ?c .
      ?c fgo:hasPattern ?p .
      GRAPH ?p { ?x rdf:type amazon:SearchCriteria } 
    }
    UNION
    {
      ?bb fgo:hasPostCondition ?c .
      ?c fgo:hasPattern ?p .
      GRAPH ?p { ?x rdf:type amazon:Item } 
    }
  }
  FILTER (?bb != <http://fast.org/screens/S1>) 
  FILTER (?bb != <http://fast.org/screens/S2>) 
}
\end{verbatim}
\label{lis:sparql_find}
\end{listing}

However, you might have realised that the pre-condition \emph{c3} would be satisfied whether the screen \emph{s1} were 
reachable, hence it will removed those pre-conditions that would be satisfied by the current elements being used in
the composition. The removal of unnecessary pre-conditions improves the performance of the algorithm, and it leads to find
better results for the user, since the focus is given to make satisfy the pre-conditions which could not be satisfied
by any of the elements of the current composition.

\subsubsection{Enhanced discovery: search tree planning}
\label{sssec:planning}

In artiﬁcial intelligence, the term \emph{planning} originally meant a search for a sequence of logical operators or actions
that transform an initial world state into a desired goal state. Presently, it also includes many decision-theoretic ideas,
imperfect state information, and game-theoretic equilibria.

This paper applies the concept of planning to the design of building blocks. Back to the screen-flow design, the goal is to
make the screen-flow executable, the initial state is a certain screen which is not yet reachable, and the plans are sets of 
screens which are reachable by themselves, and accomplish the goal. The algorithm has been influenced by the ideas behing the 
heuristic search. For a certain state, i.e. the initial state which contains the screen to make reachable, a large tree of 
possible continuations is considered, in fact any screen would fit. Those screens which post-conditions do not satisfy in any
form the unsatisfied pre-conditions are discarded, reducing the branches of the tree. A branch stops growing when a screens is
already reachable (i.e. it has no pre-conditions) becoming a leaf of the tree. Once there are not screens added in a certain
step, the algorithm stops and discards the incomplete branches, in other words, those branches where the leaf is not reachable.

It is worth pointing out some of the tree structure is pre-computed beforehand. Any time a building block is inserted into the
catalogue, the algorithm is executed following two approaches: \emph{forward search} and \emph{backward search}. The forward
search approach finds the building blocks whose pre-conditions will be satisfied by the post-conditions of the new building 
block while the backward search finds the building blocks whose post-conditions will satisfy the pre-conditions of the recently
created building block. This data is stored and used for the tree or plan builder algorithm, without needing to chec the
compatibility of pre- and post-conditions for every single building block at every request, increasing highly the performance.

\subsubsection{Results ranking}
\label{sssec:ranking}

In the previous sections there have been presented two ways of building block discovery for the composition or design of new
building blocks. The ensure compatibility based on the functional specification of the building blocks, however both are leaving
out the measurement of quality of the results. In other words, it can be seen as a search engine, returning a set of results
for a certain input, and as any other single search engine, a ranking mechanism is needed.

The simple discovery algorithm, in order to rank the results, applies the following rules:
\begin{enumerate}
 \item it gives a higher position to those building blocks which satisfy the highest number of pre-conditions;
 \item it prioritise building blocks created by the same user who is querying;
 \item it adjust the rank by using the ratings given to the building blocks, and their popularity in terms of usage statistics;
 \item it weights the results according to non-functional features such as availability. This is only applied for what we call
``web service wrapper'', and it is calculated periodically by invoking the wrapped web services generating an up-time rate.
\end{enumerate}

For the planning case, the objective is not only to produce a plan but also to satisfy user-specified preferences, or what is
known as \emph{preference-based planning}. The ranking algorithm:
\begin{enumerate}
 \item minimizes the size of the plans, after removing the elements of the plan which are already in the canvas, so it gives
priority to the elements the user has already inserted,
 \item adjust the rank by using the rules 2, 3 and 4 used in the ranking algorithm of simple discovery.
\end{enumerate}