### 2010-10-25 trunk/papers/webist2011/authors.tex
added some changes changes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1028 d8b0c59b-6544-0410-94d9-83a5658a058e

%\author{{\textmd {\textbf }}\\
%  \fontsize{11}{13}\selectfont
%  Ismael Rivera, Knud Hinnerk M\"oller\\
%  \fontsize{9}{11}\selectfont
%  \textit{DERI, National University of Ireland, Galway}\\
%  \fontsize{9}{11}\selectfont
%  \textit{ismael.rivera@deri.org, knud.moeller@deri.org}\\[12pt]
  %
%  \fontsize{11}{13}\selectfont
%  Albert Z\"undorf\\
%  \fontsize{9}{11}\selectfont
%  \textit{University of Kassel, Germany}\\
%  \fontsize{9}{11}\selectfont
%  \textit{zuendorf@cs.uni-kassel.de}\\[12pt]
  %
  % repeat previous group of commands here for further authors
%}