### 2010-10-25 trunk/papers/webist2011/paper_ssec_discovery.tex
added some changes changes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1033 d8b0c59b-6544-0410-94d9-83a5658a058e

%!TEX root = paper.tex

\subsection{Discovery mechanisms} % (fold)
\label{ssec:discovery}

The platform presented in this paper is intended to support web service discovery and recommendation capabilities.
As explained in detail in Section~\ref{ssec:ontology}, there are different types of building blocks which can be modelled
using the platform, leading to a necessity for different kinds of discovery or recommendation mechanisms, depending
on the level of composition to deal with: screen-flow or screen. In the screen-flow composition, the smallest
functional unit is the screen, meaning that a screen-flow is composed by interconnecting a number of screens between 
each other. This composition can be applied to any sort of building block since they all share the
same structure: a set of pre- and post-conditions, which make possible the communication with other building blocks.
That said, technically the algorithms presented in this section could be used for the discovery and composition of
any type of building block stored in the platform.

At the screen-flow design level, the composition is made using a set of screens, and the pre- and post-conditions 
of the screen-flow itself used as entry and exit points of the screen-flow. A screen is only reachable when its 
pre-conditions are satisfied as well, meaning that the post-conditions of another screen, or the pre-conditions of 
the screen-flow are compatible with them.

The main goal of the discovery process is to help the user to make the screen-flow executable, in other words, to make
all the screens within a screen-flow reachable. The platform offers two mechanisms to find and recommend
screens, previously stored in the building block base (the catalogue), which will make the screen-flow executable: a
simple discovery based on pre- and postconditions, and a multi-step discovery or planning algorithm. Like any other
search engine or recommender system, the results are ranked before being presented
to the user. The ranking algorithm is covered in Section~\ref{sssec:ranking}.

\subsubsection{Simple discovery based on pre- and post-conditions}
\label{sssec:simple_discovery}

At a particular state of a composition, ie. at screen-flow design, several of the building blocks used to compose it, 
screens in the case of a screen-flow, might be unreachable, therefore there must be pre-conditions not satisfied. The
platform will assist the process by recommending building blocks which will satisfy these pre-conditions. In a nutshell,
it collects the pre-conditions of all the unreachable building blocks, used as a graph pattern (see Section~\ref{ssec:ontology})
in order to be matched against the post-conditions of any building block. In the following scenario, there are two 
screens: \emph{s1} and \emph{s2}. \emph{s1} has as a pre-condition: ``there exists a search criteria'', and as a 
post-condition: ``there exists a item''. \emph{s2} has just a pre-condition stating: ``there exists a search criteria''.
Here you can see the same example but formally defined:

\begin{listing}
\begin{verbatim}
:G1 { :s1 a fgo:Screen .      
      :s1 fgo:hasPrecondition c1 .
      :s1 fgo:hasPostcondition c2 .
      :c1 fgo:hasPattern GC1 .
      :c2 fgo:hasPattern GC2 .
      :s2 a fgo:Screen .      
      :s2 fgo:hasPrecondition c3 .
      :c3 fgo:hasPattern GC3 }
:GC1 { _:x a amazon:SearchCriteria }
:GC2 { _:x a amazon:Item }
:GC3 { _:x a amazon:SearchCriteria }
\end{verbatim}
\label{lis:discovery_example}
\end{listing}

The algorithm will construct a SPARQL query to retrieve building blocks satisfying the pre-conditions \emph{c1} and \emph{c3}. 
The query, although simplified for the sake of clarity, would look something like:

\begin{listing}
\begin{verbatim}
PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
PREFIX fgo: <http://purl.oclc.org/fast/ontology/gadget#>
PREFIX amazon:<http://aws.amazon.com/AWSECommerceService#>

SELECT DISTINCT ?bb 
WHERE { 
  ?bb rdf:type fgo:Screen . 
  {
    {
      ?bb fgo:hasPostCondition ?c .
      ?c fgo:hasPattern ?p .
      GRAPH ?p { ?x rdf:type amazon:SearchCriteria } 
    }
    UNION
    {
      ?bb fgo:hasPostCondition ?c .
      ?c fgo:hasPattern ?p .
      GRAPH ?p { ?x rdf:type amazon:Item } 
    }
  }
  FILTER (?bb != <http://fast.org/screens/S1>) 
  FILTER (?bb != <http://fast.org/screens/S2>) 
}
\end{verbatim}
\label{lis:sparql_find}
\end{listing}

However, in the example presented, it is straight-forward to spot that the pre-condition \emph{c3} would be
satisfied whether the screen \emph{s1} was reachable, being unnecessary to query the recommender algorithm 
to retrieve screens satisfying this pre-condition, hence these pre-conditions are removed for the construction
of the query reducing the complexity of the problem, hence improving the overall performance of the algorithm.
Moreover, it leads to find better results for the user, since the focus is given to make satisfy the 
pre-conditions which could not be satisfied by any of the elements of the current composition.

\subsubsection{Enhanced discovery: search tree planning}
\label{sssec:planning}

In artiﬁcial intelligence, the term \emph{planning} originally meant a search for a sequence of logical operators or actions
that transform an initial world state into a desired goal state. Presently, it also includes many decision-theoretic ideas,
imperfect state information, and game-theoretic equilibria.

This paper applies the concept of planning to the design of building blocks. Back to the screen-flow design, the goal
would be to make the screen-flow executable; the initial state would be a certain screen which is not yet reachable, 
and the plans would be sets of screens which are reachable by themselves, and accomplish the goal. 
The algorithm has been influenced by the ideas behing the heuristic search. For a certain state, i.e. the initial 
state which contains the screen to make reachable, a large tree of possible continuations is considered, in 
fact any screen would fit. Those screens which post-conditions do not satisfy in any form the unsatisfied 
pre-conditions are discarded, reducing the branches of the tree. A branch stops growing when a screens is
already reachable (i.e. it has no pre-conditions) becoming a leaf of the tree. Once there are not screens 
added in a certain step, the algorithm stops and discards the incomplete branches, in other words, 
those branches where the leaf is not reachable.

It is worth pointing out some of the tree structure is pre-computed beforehand to speed up the querying process at runtime. 
Any time a building block is inserted into the catalogue, the algorithm is executed following two approaches: 
\emph{forward search} and \emph{backward search}. The forward search approach finds the building blocks whose 
pre-conditions will be satisfied by the post-conditions of the new building block while the backward search 
finds the building blocks whose post-conditions will satisfy the pre-conditions of the recently
created building block. This data is stored and used for the tree or plan builder algorithm, without
needing to check the compatibility of pre- and post-conditions for every single building block at every 
request.

\subsubsection{Results ranking}
\label{sssec:ranking}

Previous sections presented different mechanisms to search or discover building blocks which may satisfy the user needs.
The results ensure compatibility based on the functional specification of the building blocks, however they are not
ranked, and their quality is not measured. This section explains the ranking techniques applied for the different
discovery mechanisms.

The ranking algorithm for~\ref{sssec:simple_discovery} applies the following rules:
\begin{enumerate}
 \item it gives a higher position to those building blocks which satisfy the highest number of pre-conditions;
 \item it prioritise building blocks created by the same user who is querying;
 \item it adjust the rank by using the ratings given to the building blocks, and their popularity in terms of usage statistics;
 \item it weights the results according to non-functional features such as availability. This is only applied for what we call
``web service wrapper'', and it is calculated periodically by invoking the wrapped web services generating an up-time rate.
\end{enumerate}

For the planning case, the objective is not only to produce a plan but also to satisfy user-specified preferences, or what is
known as \emph{preference-based planning}. The ranking algorithm:
\begin{enumerate}
 \item minimizes the size of the plans, after removing the elements of the plan which are already in the canvas, so it gives
priority to the elements the user has already inserted,
 \item adjust the rank by using the rules 2, 3 and 4 used from the ranking algorithm of simple discovery based on 
pre- and post-conditions.
\end{enumerate}