### 2011-01-28 trunk/papers/webist2011/abstract.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

%!TEX root = paper.tex

In B2B systems integration and web services, many companies see the advantage of increased operational efficiencies and a reduction of costs. 
In this scenario, highly qualified software developers are responsible for the integration of services with other systems.
However, this model fails when targeting the long tail of enterprise software demand, the end-users. 
Discovery and consumption of web services are difficult tasks for end-users. 
This means that potential long tail of end-users creating task-specific applications from existing services is as of yet completely untapped. 
This paper presents an approach to facilitate the discovery and consumption of business web services by end-users, closing the gap between the two. The approach includes: 
\begin{inparaenum}[(a)]
	\item a catalogue which users can browse to search for web services fitting their needs, and
	\item a method to generate ready-to-use web service wrappers to use in the catalogue.
\end{inparaenum}