### 2011-01-31 trunk/papers/webist2011/authors.tex
Siggi is co-author
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1229 d8b0c59b-6544-0410-94d9-83a5658a058e

\author{{\textmd {\textbf }}\\
 \fontsize{11}{13}\selectfont
 Ismael Rivera, Knud Hinnerk M\"oller, Siegfried Handschuh\\
 \fontsize{9}{11}\selectfont
 \textit{DERI, National University of Ireland, Galway}\\
 \fontsize{9}{11}\selectfont
 \textit{ismael.rivera@deri.org, knud.moeller@deri.org, siegfried.handschuh@deri.org}\\[12pt]
  %
 \fontsize{11}{13}\selectfont
 Albert Z\"undorf\\
 \fontsize{9}{11}\selectfont
 \textit{University of Kassel, Germany}\\
 \fontsize{9}{11}\selectfont
 \textit{zuendorf@cs.uni-kassel.de}\\[12pt]
  %
  % repeat previous group of commands here for further authors
}