### 2011-02-26 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_mediation.tex
getting final version done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1305 d8b0c59b-6544-0410-94d9-83a5658a058e

\documentclass{fast_latex}
% yay! We can use any kind of funky diacritic:
\usepackage[utf8]{inputenc}

\usepackage{lastpage}
\usepackage{setspace}
\usepackage[pdfborder={0 0 0}]{hyperref}        % turn on when latex is used (not miktec)
\usepackage{url} % LEO: urls \url{}
\usepackage{verbatim} % code and comment
\usepackage{longtable}
\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
\usepackage{lipsum}
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{array}
\usepackage{textcomp} %vertical apostrophe
\usepackage{listings} %source code listings
\usepackage{amssymb} %math symbols
\usepackage{graphicx} %image placement
\usepackage{supertabular}

\usepackage{amsmath}
\usepackage{pgf} %graph plot
\usepackage{tikz}
\usetikzlibrary{plotmarks}

\usepackage{listings} %source code listings
\usepackage{paralist}
\usepackage{booktabs}

\parindent0pt

\newcommand\deliverableNumber{D2.4.3}
\newcommand\deliverableTitle{Mediation amongst ontologies: Application to the FAST platform}
\newcommand\deliverableTitleShort{Mediation amongst ontologies}
\newcommand\workpackageNumber{2}
\newcommand\workpackageTitle{Definition of Conceptual Model}
\newcommand\authorOne{Oszkar Ambrus (NUIG)}
\newcommand\authorTwo{Knud M\"{o}ller (NUIG)}
\newcommand\authorThree{Albert Z\"{u}ndorf (UniKassel)}
\newcommand\authorFour{J\"{o}rn Friedrich Dreyer (UniKassel)}
\newcommand\authorFive{Sebasti\'{a}n Ortega (UPM)}
\newtheorem{example}{\emph{Example}}

\begin{document}
% explicit hyphenations
\hyphenation{RDF-Re-po-si-to-ry}
\hyphenation{name-space}
\hyphenation{So-cket-A-dap-ter}

%\fontfamily{tahoma}\selectfont
\def\note#1{\marginpar{\footnotesize#1}} % use this to show the notes in the document
%\def\note#1{} % use this to hide the notes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE PAGES 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\thispagestyle{empty}

%\pagenumbering{roman}
\pagenumbering{arabic}

\begin{flushright}
    \includegraphics[width=3cm]{images/FP7_logo.png}
\end{flushright}

\vspace{1cm}

%\begin{minipage}[p]{15cm}
    \begin{center}
        \includegraphics{images/FAST_logo}\\
        \vspace{1cm}
        {\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
        \vspace{0.5cm}
        {\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
        \vspace{0.5cm}
        {\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
        \vspace{4cm}
        {\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
        \vspace{0.5cm}
        {\LARGE \sffamily \textbf{\deliverableTitle}}\\
        \vspace{2cm}
        {\large \sffamily \authorOne} \\
        {\large \sffamily \authorTwo} \\
        {\large \sffamily \authorThree} \\
        {\large \sffamily \authorFour} \\
        {\large \sffamily \authorFive} \\
        \vspace{0.5cm}
        \vfill
        {\large \sffamily Date: 27/02/2011}\\
        \vspace{1cm}
        {\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

    \end{center}
%\end{minipage}

\clearpage
%%%%%%%%%%%%%%
% NEXT PAGES %
%%%%%%%%%%%%%%
\pagestyle{scrheadings}

\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
%\rohead{\small{\today}}
%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
\cofoot{\small{FAST $\bullet$ 216048 $\bullet$ \deliverableTitleShort\ $\bullet$ Page \thepage\ of \pageref{LastPage}}}
%\rofoot{\small{}}
%\rofoot{\small{\thepage}}

\newpage
%
%\section*{Authors}
%\authorOne \\
%%\authorTwo \\
%%\authorThree\\
%%\authorFour\\
%
%
%\vfill
%%\section*{Project Co-ordinator}
%%\hspace*{-2,5cm}\begin{minipage}[p]{14cm}
%%Dr. Ansgar Bernardi \\
%%German Research Center for Artificial Intelligence (DFKI) GmbH \\
%%Trippstadter Strasse 122 \\
%%D 67663 Kaiserslautern \\
%%Germany \\
%%Email: bernardi@dfki.uni-kl.de, phone: +49 631 205 3582, fax: +49 631 205 4910 \\
%%\end{minipage}
%
%
%\section*{Partners}
%Telef\'onica Investigaci\'on y Desarrollo (Telef\'onica I+D), Spain \\
%National University of Ireland, Galway (NUIG), Republic of Ireland \\
%Kassel University, Germany \\
%Cyntelix Corporation, Republic of Ireland \\
%SAP AG, Germany \\
%Universidad Polit\'ecnica de Madrid (UPM), Spain
%
%
%\vspace{0.3cm}
%\begin{footnotesize}
%Copyright: FAST Consortium 2009\\
%\end{footnotesize}

%\clearpage

\section*{Version History}

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                            \textcolor{white}{\textbf{Date}} &
                            \textcolor{white}{\textbf{Author (Partner)}} &
                            \textcolor{white}{\textbf{Change description}}\\ \hline
% 0.1 & 09.05.2009 & \authorOne & Initial content \\ \hline
% 0.4 & 30.11.2009 & \authorOne & Sections 2 and 3 \\ \hline
% 0.5 & 07.01.2010 & \authorOne & New structure, Section 4 and 5 \\ \hline
% 0.7 & 03.02.2010 & \authorTwo & Section 3.5, 3.6, updated Section 5  \\ \hline
% 0.8 & 03.02.2010 & \authorOne & Integration and preliminary finalization  \\ \hline
% 0.9 & 03.02.2010 & NUIG, UniKassel & Corrections  \\ \hline
% 1.0 & 03.02.2010 & \authorOne & Integration and finalization  \\ \hline
%0.2 & 19.02.2009 & Oszkar Ambrus (NUIG) & content moved from wiki after internal review \\ \hline
%0.1 & 07.01.2009 & Oszkar Ambrus (NUIG) & final version ready for external review \\ \hline
1.0 & 27.02.2009 & \authorOne & final version (D2.4.1) ready for external review \\ \hline
2.0 & 27.02.2010 & \authorOne & final version (D2.4.2) ready for external review \\ \hline
3.0 & 27.02.2011 & \authorTwo & final version (D2.4.3) ready for external review \\ \hline
\end{tabular}
\end{small}

\color{black}

\vfill
%{\bf Explanations of abbreviations on front page}\\
%\\
%%Nature \\
%R: Report \\
%P: Prototype \\
%R/P: Report and Prototype \\
%O: Other \\
% \\
%Dissemination level \\
%PU: Public \\
%PP: Restricted to other FP6 participants \\
%RE: Restricted to specified group \\
%CO: Confidential, only for NEPOMUK partners \\

\newpage

%%%%%%%%%%%%%%%%%%%%%
% Executive Summary %
%%%%%%%%%%%%%%%%%%%%%

\clearpage

\section*{Executive Summary}
\doublespacing
Ontology mediation is a critical problem in heterogeneous environments, since inter-operating services and agents often use different
ontologies, and the differences between them need to be determined and overcome to allow for a seamless data exchange between these parties.
This is a difficult task, with which web-based agents must cope.

In \cite{ambrus2009mediation} we reviewed current work and literature, discussing the main areas of the topic: the core problem of ontology mediation, the types of mismatches that occur between ontologies, and the three main ontology mediation approaches, namely (1) the representation of correspondences between ontologies, (2) the (semi-)automatic process of discovering these correspondences and (3) creation of a new ontology from the source ontologies.

Based on these theoretical foundations, the second~\cite{ambrus2010mediation} as well as this final iteration of the deliverable present our approach to integrating ontology matching and the definition of mapping rules into the FAST platform. We present the roles related to ontology matching in FAST, and the way ontology matching is achieved within each one of those roles. We have developed and integrated a tool to allow end users to define mappings of varying complexity and create new operator building blocks to be used in the screen design process.

\newpage

%%%%%%%%%%%%%%%%%%%%%
% Document Summary %
%%%%%%%%%%%%%%%%%%%%%

\clearpage

\section*{Document Summary}
% double spacing from here on:
\singlespacing
\begin{small}

\begin{tabular}
    %{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
    {| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
    \hline
    \textcolor{white}{\textbf{Code}} & {FP7-ICT-2007-1-216048} & {\textbf{Acronym}} & {FAST}\\ \hline
    \textcolor{white}{\textbf{Full title}} & \multicolumn{3}{l|}{Fast and Advanced Storyboard Tools}\\ \hline
    \textcolor{white}{\textbf{URL}} & \multicolumn{3}{l|}{\url{http://fast.morfeo-project.eu}}\\ \hline
    \textcolor{white}{\textbf{Project officer}} & \multicolumn{3}{l|}{Annalisa Bogliolo}\\ \hline
\end{tabular}

\vspace{0.5cm}

\begin{tabular}
    {| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.25cm}|p{1cm}|p{1cm}|p{6.32cm}|}
    \hline
    \textcolor{white}{\textbf{Deliverable}} & {\textbf{Number}} & {\deliverableNumber} & {\textbf{Name}} & {\deliverableTitle}\\ \hline
    \textcolor{white}{\textbf{Work package}} & {\textbf{Number}} & {\workpackageNumber} & {\textbf{Name}} & {\workpackageTitle}\\ \hline
\end{tabular}

\vspace{0.5cm}

\begin{tabular}
    {| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
    \hline
    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2011} & {\textbf{Submitted}} & {27/02/2011}\\ \hline
    \textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
    \textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.3 is the third iteration of the FAST ontology matching and mapping approach. The first iteration contained a general overview of the literature regarding ontology mediation, paving the way for further analysis of the scenarios arising in FAST that require ontology mediation. Based on that general overview, the second and third iteration presents the approach to implement ontology matching within FAST. We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role, including the \emph{fact transformation tool}, which has become part of the FAST tool chain.}\\ \hline
    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne,  \authorTwo, \authorThree, \authorFour, \authorFive}\\
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\
    \hline
    \textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {knud.moeller@deri.org} \\ \cline{4-5}
    \textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
    \textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment}\\ \hline
\end{tabular}
\end{small}
\newpage

%%%%%%%%%%%%%%%%%%%%%
% TABLE OF CONTENTS %
%%%%%%%%%%%%%%%%%%%%%
\doublespacing
\setcounter{tocdepth}{3}
\tableofcontents
\clearpage
% \section*{Lists of Tables and Figures}
\listoftables
\listoffigures
\cleardoublepage
%\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGINNING OF SECTIONS %
%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\rofoot{\small{Page \thepage\ of \pageref{LastPage}}}

\section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:intro}
\subsection{Goal and Scope}
FAST is a visual programming platform allowing non-technical end users to build situational mashups, employing various underlying services and generating new ones. Resources in FAST are described semantically using different ontologies and vocabularies, and can therefore be combined to what we can call ``intelligent gadgets'', that rely on semantically enriched building blocks. In the services used within the FAST environment, the ontologies involved in the creation of gadgets and building blocks come from different parties. These ontologies, though different, will sometimes cover the same domain, making ontology matching necessary as the means to reconcile differences in the various conceptualisations.

We aim to integrate a form of ontology matching into the FAST platform to allow for discrepant systems to intercommunicate --- or, more precisely, previously incompatible building blocks in our case.  We try to find a solution to these differences which range from simple naming differences to more complex data heterogeneity, behavioural differences, discrepant paradigms and other disparities. The reconciliation of these differences is called ontology matching (often mentioned in the specialised literature as ontology mediation, or ontology alignment, thus we will sometimes use these terms as well).

While we have explored ways of automatically defining mappings between different conceptualisations, our focus within FAST is less on research into new such methods. Instead, we aim to support the end user in defining such mapping rules within the FAST tool chain and integrate the process within the platform. To this end, we have developed and integrated a fact transformation tool, which is part of the FAST GVS and presented in this deliverable.

\subsubsection{Changes from Previous Version}

The following changes have been made with respect to the previous version of this deliverable, D2.4.2.~\cite{ambrus2010mediation}. A history of previous changes is kept in App.~\ref{sec:old_changes}.

\begin{itemize}
	\item In response to the previous review, we have extended and improved the text in order to better position ontology matching in FAST throughout the document.
	\item The previous D2.4.2 introduced the \emph{fact transformation tool}. At the time, this tool was an early stand-alone prototype. In the meantime, the fact tool has been properly integrated with the FAST tool chain. These improvements are reflected in both text and screenshots in Sects.~\ref{sec:fact_tool} and \ref{sec:creating_the_operator}. New content has been added to show how the output of the tool is used in screen development.
	\item The document as a whole has been revised, and minor changes (mostly, spelling, grammar and small updates to reflect latest developments) have been made throughout the deliverable.
\end{itemize}

\subsection{Structure of the Document}
This document describes the approach of integrating ontology matching into the FAST environment. Sect.~\ref{sec:overview} gives the big picture of ontology matching in fast, which the remaining three sections expand. Thus, Sect.~\ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilities of the tool, and the testing procedure with the results. The remaining part of the section deals with the \emph{fact transformation tool}, developed to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue. Sect.~\ref{sec:gadgetbuilding} shortly describes the way the matching operator is used when composing screenflows and Sect.~\ref{sec:runtime} describes how the matching operator and the defined rules are converted and used in the final, deployed gadget.

%Describe the rest.
The Appendix contains the main content of first iteration of this document, D2.4.1, laying down the theoretical foundations of ontology matching. We discuss ontologies in general in App.~\ref{sec:ontologies}, present the ontology mediation problem in App.~\ref{sec:mediationproblem} , discuss the types  of mismatches between ontologies in App.~\ref{sec:ontologymismatches} and detail the main approaches to ontology mediation in App.~\ref{sec:mediationapproaches}.

\subsection{Overall Approach for WP2 and relation to other Work Packages and Deliverables.}
Ontology matching will play an important role for the integration of different underlying resources in the user-built gadgets. 
We present our initial research in the implementation of ontology matching in this second iteration of D2.4, after having laid the theoretical ground work of ontology matching in the first iteration. The final version of the deliverable (i.e., D2.4.3) will complete the implementation of ontology mediation within the FAST platform.

Parts of this work rely on the results of D4.3.2, where a service wrapper tool is described, that will be modified to be used in matching ontologies.

\include{fast_d2_4_3_part1}

%\pagebreak
%\part{D2.4.1: Theoretical Foundations for Ontology Mediation}
%\label{part:d2.4.1}

\clearpage
\phantomsection
\appendix

\include{fast_d2_4_3_part2}

\clearpage
\phantomsection
\section{Past Changes} % (fold)
\label{sec:old_changes}

For keeping a record of past changes, this section simply lists the changes that were made between the previous two versions of this deliverable, D2.4.1~\cite{ambrus2009mediation} and D2.4.2.~\cite{ambrus2010mediation}:

The first iteration, dealing with the theoretical foundations of ontology matching, had been moved to App.~\ref{sec:background}. The preceding main body of D2.4.2 document is newly added content, based on the comments of the last review (M24). As such, it tackles design and development issues for integrating ontology matching within the FAST platform: it presents the overall strategy and deals with specific problems belonging to each phase of ontology matching in FAST.

\clearpage
\bibliographystyle{apalike}
\addcontentsline{toc}{section}{References}
\bibliography{fast_d2_4_mediation_bib}

\end{document}